Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-08-30 Ben_Manuscripts/structure_paper/Draft.tex
intro edits

  More highly selective nanoporous membranes would be extremely useful in order
  to perform complex aqueous separations with seawater and various sources of
  wastewater. Sodium chloride and boron in seawater \cite{fritzmann_state---art_2007}, and organic micropollutants
  found in municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006} are examples of the
  diverse constituents of various water sources. By efficiently
  separating contaminants from feed solutions with highly selective membranes,
  one can reduce the number of required membrane passes and post-treatment steps
  needed for a given filtration process \cite{werber_materials_2016}, thus lowering energy requirements.
  Additionally, one can extract valuable resources from the feed streams. For
  example, flowback water produced during hydraulic fracturing of shale
  formations contains dissolved species such as acetate whose extraction could be
  economically valuable \cite{dischinger_application_2017}.

@@

 % (2) existing nanoporous membranes + drawbacks

@@

  % (3) justification for why nanostructured membranes are a good idea for this
  % long-term route towards highly selective membranes are those that are nanostructured
  Nanostructured porous membrane materials have become increasingly popular for
  aqueous separations applications because they offer the ability to control pore
  architecture at the molecular scale, thereby permitting the design of
  solute-specific separation membranes \cite{humplik_nanostructured_2011},
  however their development has been limited by the ability to synthesize and
  scale various fundamentally sound technologies. Graphene sheets are atomically
  thick which results in excellent water permeability but defects during
  manufacturing severely impact selectivity \cite{cohen-tanugi_multilayer_2016}.
  Molecular dynamics (MD) simulations of carbon nanotubes show promise
  \cite{humplik_nanostructured_2011}, but synthetic techniques are unable to
  achieve scalable alignment and pore monodispersity
  \cite{hata_water-assisted_2004,maruyama_growth_2005}. Consequently, there is a
  need for a scalable nanostructured porous membrane.

@@

%MRS8: preliminary evidence has shown they may be?  It's more than a guess.
  Preliminary evidence has shown that crosslinked lyotropic liquid crystal
  (LLC) membranes may be capable of performing highly selective separations. LLCs
  are amphiphilic molecules that have the ability to self-assemble into porous
  nanostructures \cite{smith_ordered_1997} and can be crosslinked to create
  mechanically strong membrane films with pores on the order of 1 nm in diameter
  \cite{zhou_supported_2005}. Unlike most commercial NF membranes, LLC membrane
  pores are uniform in size because they are self-assembled. Since LLC membranes
  lack a pore size distribution, they inherently exhibit high selectivity due to
  their strict molecular weight cut-off (MWCO)~\cite{zhou_supported_2005}.
  Additionally, the LLC monomers examined in this paper are salts, and therefore
  lead to Donnan exclusion of ions in solution. The membrane gains a net surface
  charge when counterions from the head groups that line the pore walls escape
  into the feed solution in an effort to balance the gradients of concentration
  and electric potential \cite{donnan_theory_1995}.

@@

  The H\textsubscript{II} phase pore geometry (Figure~\ref{fig:assembly}) has a
  higher theoretical capacity for transport than the Q\textsubscript{I} phase.
  The H\textsubscript{II} phase forms at room temperature in the presence of
  c.a.~10 wt\% water and consists of hexagonally packed, hydrophilic pore
  columns\cite{smith_ordered_1997}. In the absence of water, neat monomer will
  form the same hexagonal columnar structure which, in literature, has been
  referred to as the Col\textsubscript{h} thermotropic
  phase\cite{feng_scalable_2014}. 
  % BJC6: maybe not a necessary detail
%  The most
%  promising Q\textsubscript{I} phase for membrane applications forms at 70\degree
%  C when monomer is mixed with c.a.~20 wt\% glycerol\cite{carter_glycerol-based_2012}.
  Q\textsubscript{I} phase membranes consist of a tortuous network of three
  dimensionally interconnected pores that prevent optimal through-plane
  transport. In contrast, the densely packed, non-tortuous and uniform sized
  pores of H\textsubscript{II} phase membranes represent the ideal geometry for
  achieving high solute flux\cite{matyka_tortuosity-porosity_2008}.  However, the
  hexagonally packed liquid crystalline domains of the H\textsubscript{II} phase
  generally form isotropically oriented and therefore mutually unaligned domains
  which are detrimental to membrane permeability. This domain scale misalignment
  has inhibited further development of the this technology, and research efforts
  were focused on the Q\textsubscript{I} phase, whose geometry does not require
  alignment~\cite{dischinger_papers}.

@@

>>>>

  More highly selective nanoporous membranes would be extremely useful in order
  to perform complex aqueous separations with seawater and various sources of
  wastewater. Sodium chloride and boron in seawater
  \cite{fritzmann_state---art_2007}, and organic micropollutants found in
  municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006} are
  examples of the diverse constituents of these water sources. By efficiently
  separating contaminants from feed solutions with highly selective membranes,
  one can reduce the number of required membrane passes and post-treatment steps
  needed for a given filtration process \cite{werber_materials_2016}, thus
  lowering energy requirements. Additionally, one can extract valuable resources
  from the feed streams. For example, flowback water produced during hydraulic
  fracturing of shale formations contains dissolved species such as acetate whose
  extraction has economic value \cite{dischinger_application_2017}.

@@

 % BJC14: (2) existing nanoporous membranes + drawbacks

@@

  % BJC14: (3) justification for why nanostructured membranes are a good idea for this
  % long-term route towards highly selective membranes are those that are nanostructured
  Nanostructured porous membrane materials have become increasingly popular for
  aqueous separations applications because they offer the ability to control pore
  architecture at the molecular scale, thereby permitting the design of
  solute-specific separation membranes \cite{humplik_nanostructured_2011},
  however their development has been limited by the ability to synthesize and
  scale various fundamentally sound technologies. Graphene sheets are atomically
  thick which results in excellent water permeability but defects during
  manufacturing severely impact selectivity \cite{cohen-tanugi_multilayer_2016}.
  Molecular dynamics (MD) simulations of carbon nanotubes show promise
  \cite{humplik_nanostructured_2011}, but synthetic techniques are unable to
  achieve scalable alignment and pore monodispersity
  \cite{hata_water-assisted_2004,maruyama_growth_2005}. Consequently, there is a
  need for a scalable nanostructured porous membrane.

@@

%MRS8: preliminary evidence has shown they may be?  It's more than a guess.
  Preliminary evidence has shown that crosslinked lyotropic liquid crystal
  (LLC) membranes are scalable and may be capable of performing highly selective
  separations. LLCs are amphiphilic molecules that have the ability to
  self-assemble into porous nanostructures \cite{smith_ordered_1997} and can be
  crosslinked to create mechanically strong membrane films with pores on the
  order of 1 nm in diameter \cite{zhou_supported_2005}. Unlike most commercial NF
  membranes, LLC membrane pores are uniform in size because they are
  self-assembled. Since LLC membranes lack a pore size distribution, they
  inherently exhibit high selectivity due to their strict molecular weight
  cut-off (MWCO)~\cite{zhou_supported_2005}. Additionally, the LLC monomers
  examined in this paper are salts, and therefore lead to Donnan exclusion of
  ions in solution. The membrane gains a net surface charge when counterions from
  the head groups that line the pore walls escape into the feed solution in an
  effort to balance the gradients of concentration and electric potential
  \cite{donnan_theory_1995}.

@@

  The H\textsubscript{II} phase pore geometry (Figure~\ref{fig:assembly}) has a
  higher theoretical capacity for transport than the Q\textsubscript{I} phase.
  The H\textsubscript{II} phase forms at room temperature in the presence of
  c.a.~10 wt\% water and consists of hexagonally packed, hydrophilic pore
  columns\cite{smith_ordered_1997}. In the absence of water, neat monomer will
  form the same hexagonal columnar structure which, in literature, has been
  referred to as the Col\textsubscript{h} thermotropic
  phase\cite{feng_scalable_2014}. 
  % BJC6: maybe not a necessary detail
%  The most
%  promising Q\textsubscript{I} phase for membrane applications forms at 70\degree
%  C when monomer is mixed with c.a.~20 wt\% glycerol\cite{carter_glycerol-based_2012}.
  Q\textsubscript{I} phase membranes consist of a tortuous network of three
  dimensionally interconnected pores that prevent optimal through-plane
  transport. In contrast, the densely packed, non-tortuous and uniform sized
  pores of H\textsubscript{II} phase membranes represent the ideal geometry for
  achieving high solute flux\cite{matyka_tortuosity-porosity_2008}.  However, the
  hexagonally packed liquid crystalline domains of the H\textsubscript{II} phase
  generally form isotropically oriented and therefore mutually unaligned domains
  which are detrimental to membrane permeability. This domain scale misalignment
  have inhibited further development of the this technology, and research efforts
  were focused on the Q\textsubscript{I} phase, whose geometry does not require
  alignment~\cite{dischinger_papers}.

@@

  %BJC14: nothing new after this point
